### 2019-06-25 2-Chapters/Conclusion/conclusion.tex
Lots of typos

This manuscript was organized in two parts, Part~\ref{part:Introduction} and Part~\ref{part:MABIOT}.
It presented the problems we considered for my PhD research, and we discussed about the main contributions produced during the last three years.

% \begin{itemize}
    % \item
The first half, Part~\ref{part:Introduction}, started by motivating our work and situating the context of this research.
% Chapter 2
In Chapter~\ref{chapter:2}, we presented the multi-armed bandit models, the hypotheses we usually make, and the most important algorithms that solve different kinds of MAB problems.
% Chapter 3
Then in Chapter~\ref{chapter:3} we discussed in details about our open-source Python library, SMPyBandits, developed to run numerical simulations of MAB problems.
%
Finally, we presented in Chapter~\ref{chapter:25} a first mathematical contribution, our algorithm \Aggr, as an efficient answer to the online algorithm selection problem.

    % \item
The second half of this thesis, Part~\ref{part:MABIOT}, presented our main contributions.
% Chapter 4
In Chapter~\ref{chapter:4}, we first proposed some models for realistic IoT wireless networks, in discrete times and with no central coordination between end-devices.
In two different models, with or without retransmissions, we showed that end-devices can learn independently to increase their successful transmission rates in the network, by using low-cost MAB algorithms. We also presented a proof-of-concept that implements such network on real radio hardware and validates our model and proposed algorithms.
% Chapter 5
Because the models studied for realistic IoT networks were found to be intractable to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in an unpredictable environment,
in Chapter~\ref{chapter:5}, we studied a weaker but still interesting model.
We presented different variants of the multi-players MAB model, and we proposed the state-of-the-art algorithm for the variant with collision information (\MCTopM-\klUCB).
% Chapter 6
Finally, in Chapter~\ref{chapter:6}, we removed the stationary hypothesis on the single-player MAB problem, and we analyzed a new actively adaptive algorithm, for which we also proved state-of-the-art results (\GLRklUCB).
%
For the last two chapters, our proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret bounds obtained in our theoretical analyses.
% \end{itemize}

@@

Then we presented our Python library SMPyBandits \cite{SMPyBandits,SMPyBanditsJMLR}, by detailing
its purpose and qualities, its organization, and an overview of its usage.
%
SMPyBandits allows any researcher to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license, and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
% Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.

We detailed how SMPyBandits is implementing arms, problems, algorithms, and how the library uses these components to implement a simulation loop, with various visualizations being performed after a simulation.
Even if SMPyBandits is still restricted to the finite-arm case, it supports a wide range of arm distributions (as well as rested or restless Markov models).
Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature.
% We indeed followed actively the state-of-the-art research, since December $2016$ to June $2019$.
More than $65$ algorithms (or variants) are implemented for the single-player case, $5$ for the experts aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary case.
The entire codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
%
When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.
%
Our library is used to run numerical simulations on Chapters~\ref{chapter:2}~ \ref{chapter:5} and~\ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.

@@

First in Section~\ref{sec:4:firstModel}, we proposed an evaluation of the performance of applying MAB learning algorithms in IoT networks,
with a focus on the convergence of algorithms, in terms of successful transmission rates.
We restrict to the case of a protocol slotted in both time and frequency, and we assume a certain clock pre-agreement and time synchronisation between devices and the base station.
We studied the behavior of the network when the proportion of intelligent dynamic devices changes.
Concretely, increasing this probability allows to insert more devices in the same network, while maintaining a good Quality of Service.
Similarly, if the number of devices remain constant, increasing the successful transmission rate directly improves the network QoS.
%  as the base station receives more accurately the data sent by the end-devices since less transmissions failed.
We show that both \UCB{} and Thompson sampling algorithms have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which follow a random activation process.
%
This is both surprising and encouraging, it shows that applying bandit algorithms tailored for a stochastic model is still useful in broader settings.
The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each other, and \emph{do not} know the system parameters.

We presented in Section~\ref{sec:4:gnuradio} a proof-of-concept, demonstrated in $2018$ at the ICT conference \cite{Besson2018ICT}, and further detailed in the companion paper \cite{Besson2019WCNC}.
We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
Results obtained in practice were discussed, to highlight the interest of using learning algorithms for radio online optimization problem, and especially multi-armed bandit learning algorithms.
%
By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using onlt the acknowledgement (\Ack) feedback sent from the base station.

Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, again under the hypothesis of a protocol slotted in time and frequency and with perfect synchronisation.
% , in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
If the priority is the Quality of Service (QoS), for instance with renewable energy capabilities, this second model is more appropriate.
We presented and evaluated several heuristics that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission (first-stage), and different proposals based on \UCB{} for the retransmissions upon collisions (second-stage).
The main novelty of this model is to also learn the most efficient way to retransmit packets upon radio collision, by also using a multi-armed bandit algorithm for this second-stage optimization problem.
%
We showed that incorporating learning for the transmission is needed to achieve optimal performance, with significant gain in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
Our empirical simulations show that each of our proposed heuristic outperforms a naive random access scheme.
Surprisingly, the main take-away message is that a simple \UCB{} learning approach (that simply retransmits in the same channel), turns out to perform almost as well as more complicated heuristics.

Thus in Chapter~\ref{chapter:4}, we proposed two models of Internet of Things (IoT) networks, composed of many independent IoT end-devices,
that can use low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
%
Decentralized RL for IoT lets the devices use acknowledgements sent by their Base Station as a reward, instead of sensing feedback like for OSA.
We consider many independent ``dynamic'' devices, communicating with a small probability at every instant.
The first model (without retransmission of packets),
is interesting for its simplicity, and because considering no retransmission can improve the battery life of the IoT devices, at the cost of a lower Quality of Service (QoS).
%
However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission (\eg, $10$ times).
%
The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.

@@

The second chapter of Part~\ref{part:MABIOT} of this manuscript studied Multi-Players Multi-Arm Bandits models.
%
We present three variants of this model,
with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
For the two easiest models --with sensing--, our theoretical contribution improves the state-of-the-art upper bounds on the regret. In the absence of sensing, we also provide some motivation for the practical use of the \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
%
We also reviewed various variants of this model, and for some interesting variants we discussed the related literature, which has proven to be very active in the last two years.
For some of these variants, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt \MCTopM{} to other settings.
For example, it assumes to know the number of players $M$ before-hand, but we illustrated that previously introduced technique to estimate $M$ can also be applied to our proposal and give satisfactory empirical performances.
% Further works would be required to adapt the theoretical analysis to these various extensions.

@@

In the last chapter, we were interested in relaxing the stationary hypothesis made in Chapter~\ref{chapter:2}.
%
We first describe the break-point detection problem, that we solve by analyzing at finite-time the properties of the Bernoulli GLR change-point detector. This sequential break-point detection test can be applied for bounded rewards, and we bound its false alarm probability as well as its detection delay.
For the piece-wise stationary bandit model,
we proposed a new efficient algorithm, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector.
This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but without any other prior knowledge on the problem, unlike \CUSUMUCB{} and \MUCB{} that require to know a lower bound on the smallest magnitude of a change.
We also gave numerical evidence of the efficiency of our proposal, which performs usually much better than all the passively adaptive approaches as well as better or comparably to the previous actively adaptive ones.

@@

As detailed in Chapter~\ref{chapter:2}, we focused in this thesis on finite-arms and one-dimensional bandit problems,
and thus two possible directions of future research could be to extend our works
to MAB models with either multidimensional rewards, like contextual bandits, or infinite arms, like Lipschitz bandits.
The hypotheses we made were motivated by the target applications, but these other models have also started to be used for cognitive radio applications.

@@

A few tasks are left on our library SMPyBandits, a first one could be to implement new variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
%
A second interesting task could be to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
or to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
%
A third task that I would have liked to complete is to interface the library with a web-based interactive demonstration, in order to allow anyone to launch simulations without any knowledge of programming.
It is already possible to reproduce some of the experiments presented in this thesis, by using the Jupyter notebooks \cite{jupyter}, hosted on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}, which can be used locally if you install the library, but can also be used on cloud platforms, such as
Binder\footnote{~You can try on \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} from your browser.}
or Google Colab\footnote{~See \href{https://colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}{\texttt{colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}}.}.

@@

The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices have different transmission pattern, and less easily with one probability per device. Also, other emission patterns could be considered, instead of a Bernoulli process for each user.
In this whole Chapter~\ref{chapter:4}, we prefer to consider that all devices have the same activation probability, to keep the notation as simple as possible.
%
Moreover, for sake of simplicity we supposed that all devices use the same standard.
Future works could consider more realistic interference scenarios and IoT networks, with, \eg, non-slotted time, more than one base station etc.

@@

Our study on multi-players bandits suggests several further research directions.
First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-players model with sensing information. So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality. However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus a promising future work is to identify some notion of \emph{minimal number of collisions}.
% Similarly to what was done very recently in \cite{wang2019distributed} for the communications between players who collaborate with each other, it would be interesting to characterize the number of collisions needed to achieve logarithmic regret.

@@

We believe that our new proof technique could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous work), that would require only a few ``meaningful'' changes to be detected. This promising research direction is left for future work,  but the hope is that the regret would be expressed in term of this number of meaningful changes instead of $\Upsilon_T$. We shall also investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.
We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as \GLRklUCB, could lead to an interesting class of algorithms, as suggested by \cite{Maillard2018GLR}.

>>>>

This manuscript was organized in two parts, Part~\ref{part:Introduction} and Part~\ref{part:MABIOT}.
%
The main problem we studied was \emph{``Can we adapt the decision making tools already successfully applied to Cognitive Radio for Opportunistic Spectrum Access to the specific needs of CR for the (future) Internet of Things networks?''}
%
This document detailed the historical, scientific and technical contexts of the problems we considered for our research, and presented the main contributions produced during the last three years.

% \begin{itemize}
    % \item
The first half, Part~\ref{part:Introduction}, started by motivating our work and situating the context of this research.
% Chapter 2
In Chapter~\ref{chapter:2}, we presented the multi-armed bandit models, the hypotheses we usually make, and the most important algorithms that solve different kinds of MAB problems.
% Chapter 3
Then in Chapter~\ref{chapter:3} we discussed in details about our open-source Python library, SMPyBandits, developed to run numerical simulations of MAB problems.
%
Finally, we presented in Chapter~\ref{chapter:25} a first mathematical contribution, our algorithm \Aggr, as an empirically efficient answer to the online algorithm selection problem.

    % \item
The second half of this thesis, Part~\ref{part:MABIOT}, presented our main contributions.
% Chapter 4
In Chapter~\ref{chapter:4}, we first proposed some models for realistic IoT wireless networks, in discrete times and with no central coordination between end-devices.
In two different models, with or without retransmissions, we showed that end-devices can learn independently to increase their successful transmission rates in the network, by using low-cost MAB algorithms.
We also presented a proof-of-concept that implements such network on real radio hardware and validates our model and proposed algorithms.
% Chapter 5
Because the models proposed for realistic IoT networks were found to be intractable to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in an unpredictable environment,
in Chapter~\ref{chapter:5}, we studied a weaker but still interesting model.
We presented different variants of the multi-players MAB model, and we proposed the state-of-the-art algorithm for the variant with collision information (\MCTopM-\klUCB).
% Chapter 6
Finally, in Chapter~\ref{chapter:6}, we relaxed the stationary hypothesis on the single-player MAB problem, and we analyzed a new actively adaptive algorithm, for which we also proved state-of-the-art results (\GLRklUCB).
%
For the last two chapters, our proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret bounds obtained in our theoretical analyses.
% \end{itemize}

@@

Then we presented our Python library SMPyBandits \cite{SMPyBandits,SMPyBanditsJMLR}, by detailing
its purpose and qualities, its organization, and an overview of its usage.
%
Our library allows any researcher to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license (MIT License), and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
% Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.

We detailed how SMPyBandits is implementing arms, problems, algorithms, and how the library uses these components to implement a simulation loop, with various visualizations being performed after a simulation.
Even if SMPyBandits is still restricted to the finite-arm case, it supports a wide range of arm distributions (and rested/restless Markov models).
Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
One of the main qualities of our library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature.
% We indeed followed actively the state-of-the-art research, since December $2016$ to June $2019$.
More than $65$ algorithms (or variants) are implemented for the single-player case, $5$ for the experts aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary case.
The entire codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
%
When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.
%
Our library is used to run numerical simulations on Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.

@@

First in Section~\ref{sec:4:firstModel}, we proposed an evaluation of the performance of applying MAB learning algorithms in IoT networks,
with a focus on the convergence of algorithms, in terms of successful transmission rates.
We restrict to the case of a protocol slotted in both time and frequency, and we assume a certain clock pre-agreement and time synchronisation between devices and the base station.
We studied the behavior of the network when the proportion of intelligent dynamic devices changes.
Concretely, increasing this probability allows to insert more devices in the same network, while maintaining a good Quality of Service.
Another way of viewing this result is the opposite point-of-view: if the number of devices remains constant, increasing the successful transmission rate directly improves the network QoS.
%  as the base station receives more accurately the data sent by the end-devices since less transmissions failed.
We show that both \UCB{} and Thompson sampling algorithms have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which all follow a random activation process.
%
This was both surprising and encouraging, as it shows that applying bandit algorithms tailored for a stochastic model is still useful in broader settings.
The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each other, and \emph{do not} know the system parameters.

We presented in Section~\ref{sec:4:gnuradio} a proof-of-concept, demonstrated in $2018$ at the ICT conference \cite{Besson2018ICT}, and further detailed in the companion paper \cite{Besson2019WCNC}.
We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
Results obtained in practice were discussed, to highlight the interest of using learning algorithms for radio online optimization problem, and especially multi-armed bandit learning algorithms.
%
By using such low-cost algorithms, we confirmed empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using only the acknowledgement (\Ack) feedback sent from the base station.

Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of our model of Low-Power Wide-Area (LPWA) networks based on an ALOHA protocol, again under the hypothesis of a protocol slotted in time and frequency, and assuming a perfect synchronisation.
% , in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
If the priority is the Quality of Service (QoS), for instance with renewable energy capabilities, this second model is more appropriate.
We presented and evaluated several heuristics that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for the first transmission (first-stage), and different proposals based on \UCB{} for the retransmissions upon collisions (second-stage).
The main novelty of this model is to also learn the most efficient way to retransmit packets upon radio collision, by also using a multi-armed bandit algorithm for this second-stage optimization problem.
%
We showed that incorporating learning for the transmission is needed to achieve optimal performance, with significant gain in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
Our empirical simulations show that each of our proposed heuristic outperforms a naive random access scheme.
Surprisingly, the simulations showed that a simple \UCB{} learning approach, that simply retransmits in the same channel, turns out to perform almost as well as more complicated heuristics, that can change the channel used for retransmissions.

Thus in Chapter~\ref{chapter:4}, we proposed two models of Internet of Things (IoT) networks, composed of many independent IoT end-devices,
that can all use low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
%
Decentralized RL for IoT lets the devices use the acknowledgement sent by their Base Station as a reward, instead of requiring some sensing capacity, like in the OSA case.
We consider many independent ``dynamic'' devices, communicating with a small probability at every instant.
The first model (without retransmission of packets),
is interesting for its simplicity, and because considering no retransmission can improve the battery life of the IoT devices.
% , at the cost of a lower Quality of Service (QoS).
%
However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission (\eg, $10$ times).
%
The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.

@@

The second chapter of Part~\ref{part:MABIOT} of this manuscript studied Multi-Players Multi-Arm Bandits models.
%
We present three variants of this model,
with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
For the two easiest models --with sensing, our theoretical contribution improves the state-of-the-art upper bounds on the regret.
In the absence of sensing, we also provide some motivation for the practical use of the \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
\Selfish{} corresponds to the heuristic used in the first model studied in Chapter~\ref{chapter:4}.
%
We also reviewed various variants of this model, and for some interesting variants we discussed the related literature, which has proven to be very active in the last two years.
For some of these variants, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt \MCTopM{} to other settings.
For example, it assumes to know the number of players $M$ before-hand, but we illustrated that a previously introduced technique to estimate $M$ can also be applied to our proposal and give satisfactory empirical performances.
% Further works would be required to adapt the theoretical analysis to these various extensions.

@@

In the last chapter, we were interested in relaxing the stationary hypothesis made in Chapter~\ref{chapter:2}.
%
We first describe the break-point (or change-point) detection problem, that we solve by analyzing at finite-time the properties of the Bernoulli GLR change-point detector.
This sequential break-point detection test can be applied for bounded rewards, and we bound its false alarm probability as well as its detection delay.
For the piece-wise stationary bandit model,
we proposed a new efficient algorithm, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector.
This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but without any other prior knowledge on the problem, unlike \CUSUMUCB{} and \MUCB{} that require to know a lower bound on the smallest magnitude of a change.
We also gave numerical evidence of the efficiency of our proposal, which performs usually much better than all the passively adaptive approaches as well as better or comparably to the previous actively adaptive ones.

@@

As detailed in Chapter~\ref{chapter:2}, we focused in this thesis on MAB problems with a finite number of arms, and binary, bounded or real-valued rewards.
Thus two possible directions of future research could be to extend our works
to MAB models with either multidimensional rewards, like contextual bandits, or infinite arms, like Lipschitz bandits.
The hypotheses we made were motivated by the target applications, but these other models have also started to be used for cognitive radio applications.

@@

A few tasks are left on our library SMPyBandits, a first one could be to implement new variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
%
A second interesting task could be to extend the library for problems with non-finite arms, such as linear or contextual bandits (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
or to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
%
A third task that I would have liked to complete is to interface the library with a web-based interactive demonstration, in order to allow anyone to launch simulations without any knowledge of programming.
It is already possible to reproduce some of the experiments presented in this thesis, by using the Jupyter notebooks \cite{jupyter}, hosted on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}, which can be used locally if you install the library, but can also be used on cloud platforms, such as
Binder\footnote{~You can try on \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} from your browser.}
or Google Colab\footnote{~See \href{https://colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}{\texttt{colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}}.}.

@@

The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices have different transmission pattern, and less easily with one probability per device.
Also, other emission patterns could be considered, instead of a Bernoulli process for each user.
In this whole Chapter~\ref{chapter:4}, we prefer to consider that all devices have the same activation probability, to keep the notation as simple as possible.
%
Moreover, for the sake of simplicity we supposed that all devices use the same standard.
Future works could consider more realistic interference scenarios and IoT networks, with, \eg, non-slotted time, more than one base station etc.

@@

Our study on multi-players bandits suggests several further research directions.
First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-players model with sensing information.
So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus a promising future work is to identify some notion of \emph{minimal number of collisions}, like a recent work \cite{wang2019distributed} studied the minimal amount of communication needed to achieve logarithmic regret, in a similar model that authorizes direct communications between players.
% Similarly to what was done very recently in \cite{wang2019distributed} for the communications between players who collaborate with each other, it would be interesting to characterize the number of collisions needed to achieve logarithmic regret.

@@

We believe that our new proof technique could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous works), that would require only a few ``meaningful'' changes to be detected.
This promising research direction is left for future work,  but the hope is that the regret would be expressed in term of this number of meaningful changes instead of the number of break-points $\Upsilon_T$.
We shall also investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.
We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as \GLRklUCB, could lead to an interesting class of more efficient algorithms, as suggested by \cite{Maillard2018GLR}.
